Note on: Out-of-Sample Error for Laplacian Eigenmaps

### 2021-03-02 notes/weekly_notes/12.24.20.NotesForWeek.tex
More progress made on the draft of Laplacian Eigenmaps paper.

For a kernel function $\eta(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $\varepsilon > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{\varepsilon,Q}$ is given by
\begin{equation*}
\bigl(T_{\varepsilon,Q}f)(x) := \frac{1}{d_Q(x)} \int_{\Omega} f(z)\eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dQ(z)
\end{equation*}
where $d_Q(x) = \int_{\Omega} \eta\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$. We denote $T_{\varepsilon,P_n}$ by $T_{\varepsilon,n}$, and $d_{P_n}$ by $d_n$. For convenience, we will assume that $\eta$ is normalized, $\int_{0}^{\infty} \eta(\|z\|) \,dz = 1$, and we will denote $K(\cdot) = (\eta(\cdot))^2$.

The estimate we will consider is a kernel smoother passed over $\wh{\theta}$,
\begin{equation}
\label{eqn:kernel_smoother_laplacian_eigenmaps}
\wh{f}(x) := T_{\varepsilon,n}\wh{\theta}_{1:n}(x).
\end{equation}
In order to relate the out-of-sample error $\|\wh{f} - f^{\ast}\|_P$ to the in-sample error $\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n$, we will assume the following.
\begin{enumerate}[label=(A\arabic*)]
	\item
	\label{asmp:kernel}
	The kernel function $\eta$ is supported on a subset of $[0,1]$, with $\eta(1/2) > 0$. Additionally, $\eta$ is \textcolor{red}{Lipschitz} on its support.
	\item 
	\label{asmp:domain} There exists $\alpha > 0$ such that for every $x \in \mc{X}$, $d_{\nu}(x) \geq \alpha \cdot \varepsilon^d$.
	\item 
	\label{asmp:density}The distribution $P$ admits a density $p$ with respect to Lebesgue measure. The density $p \in C^1(\mc{X})$, and is bounded away from $0$ and $\infty$ for all $x \in \mc{X}$,
	\begin{equation*}
	0 < p_{\min} \leq p(x) < p_{\max} < \infty.
	\end{equation*}
\end{enumerate}
Under Assumption~\ref{asmp:kernel}, we automatically have that $d_\nu(x) = \varepsilon^d$ for all $x \in \mc{X}_{\varepsilon}$. Assumption~\ref{asmp:domain} is thus a condition on the boundary $\partial \Omega$, and is implied for all sufficiently small $\varepsilon$ if the boundary is Lipschitz. 
\begin{lemma}
	\label{lem:kernel_smoothing_laplacian_eigenmaps}
	Suppose $f^{\ast} \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$. Then with probability at least $1 - \delta - C\varepsilon^d\exp\{-Cn\varepsilon^d\}$, it holds that
	\begin{equation*}
	\|\wh{f} - f^{\ast}\|_P^2 \leq C\biggl(\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n^2 + \frac{1}{\delta} \cdot \frac{\varepsilon^2}{n\varepsilon^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2\biggr)
	\end{equation*}
\end{lemma}
Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data (on the order of $1/(n\varepsilon^d)$) by a factor of $\varepsilon^2$. The bias term, on the other hand, is standard, and in particular the following estimate of the bias can be obtained. 
\begin{lemma}
	Suppose $f^{\ast} \in H_0^{s}(\Omega)$, $p \in C^{r}(\Omega)$ for $r = s - 1$, and that $\eta$ is an order-$s$ kernel. Then
	\begin{equation*}
	\|T_{P,\varepsilon}f^{\ast} - f^{\ast}\|_P^2 \leq C \varepsilon^{2s}.
	\end{equation*}
\end{lemma}
\begin{proof}
	\textcolor{red}{(TODO)}
\end{proof}

@@

\begin{proof}[Proof of Lemma~\ref{lem:kernel_smoothing_laplacian_eigenmaps}]
	By the triangle inequality, 
	\begin{equation*}
	\|\wh{f} - f^{\ast}\|_P \leq \|\wh{f} - T_{\varepsilon,n}f^{\ast}_{1:n}\|_P + \|T_{\varepsilon,n}f^{\ast}_{1:n} - f^{\ast}\|_P.
	\end{equation*}
	We analyze each of the two terms in the above expression separately. We will show that
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_0}
	\|\wh{f} - T_{\varepsilon,n}f^{\ast}_{1:n}\|_P \leq \frac{1}{\alpha}\bigl(1 + o_P(1)\bigr) \cdot \|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n
	\end{equation}
	and 
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_0.5}
	\|T_{\varepsilon,n}f^{\ast}_{1:n} - f^{\ast}\|_P^2 \leq \frac{1}{\delta}\frac{\varepsilon^2}{n\varepsilon^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2,
	\end{equation}
	the latter inequality holding with probability $1 - \delta$.

@@

	The left hand side of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is equal to $\|T_{\varepsilon,n}(\wh{\theta}_{1:n} - f^{\ast}_{1:n})\|_P$, and~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is implied by showing that the operator $T_{\varepsilon,n}: L^2(X) \to L^2(\Omega)$ is (up to constant) a contraction, meaning that for any $u \in L^2(X)$,
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_1}
	\|T_{\varepsilon,n}u\|_P^2 \leq \frac{1}{\alpha}\bigl(1 + o_P(1)\bigr) \|u\|_n^2.
	\end{equation}
	We begin with a pointwise estimate, which follows from Jensen's inequality:
	\begin{equation*}
	\Bigl(\bigl(T_{n,\varepsilon}u\bigr)(x)\Bigr)^2 \leq \frac{1}{d_n(x) n} \sum_{i = 1}^{n} (u_i)^2 \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
	\end{equation*}
	(Note that it is here that we have used the positivity of $\eta(\cdot)$.) Integrating with respect to $P$ and exchanging sum with integral yields an upper bound on the $L^2(P)$ norm of $T_{n,\varepsilon}u$,
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_2}
	\|T_{n,\varepsilon}u\|_P^2 \leq \sum_{i = 1}^{n} u_i^2 \Bigl(\int_{\Omega} \frac{\eta(\|X_i - x\|/\varepsilon)}{n d_n(x)} p(x) \,dx \Bigr)
	\end{equation}
	The degree estimate $d_n(x)$ is quite close to $p(x)$. As we show in Lemma~\ref{lem:measure_of_ball_estimate}, if $\varepsilon = \omega((\log n/n)^{1/d})$ then the empirical measure $d_n(x)$ concentrates around $d_P(x)$ uniformly over $x \in \Omega$, 
	\begin{equation*}
	\sup_{x \in \Omega}|d_n(x) - d_P(x)| = o_P(\varepsilon^d);
	\end{equation*}
	moreover, it follows from Lipschitz continuity of $p$ that for any $x \in \Omega$,
	\begin{equation*}
	d_P(x) = \int_{\Omega} \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) p(x') \,dx' = p(x) d_{\nu}(x) + O(\varepsilon^{d + 1});
	\end{equation*}
	and thus by the triangle inequality $\sup_{x \in \Omega}|d_n(x) - d_\nu(x) p(x)| = o_P(\varepsilon^d)$.

@@

	Now, suppose $\eta(\cdot)$ can take on negative values. In this case, we have $\eta(\cdot) = \mathrm{sgn}(\eta(\cdot))K^{1/2}(\cdot)$. Let $d_n(x;K^{1/2})$ be the empirical degree with respect to kernel function $K^{1/2}$. Then
	\begin{align*}
	T_{n,\varepsilon}(\wh{\theta} - \varepsilon)(x) & = \frac{1}{d_n(x)n} \sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f^{\ast}(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr) \\
	& = \frac{d_n(x;K^{1/2})}{d_n(x)}\frac{1}{d_n(x;K^{1/2})n} \sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f^{\ast}(X_i)\bigr) \mathrm{sgn}\biggl(\eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr)\biggr) K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
	\end{align*}
	Taking the square and applying Jensen's inequality, we deduce that
	\begin{align*}
	\bigl(T_{n,\varepsilon}(\wh{\theta} - \varepsilon)(x)\bigr)^2 & \leq \biggl(\frac{d_n(x;K^{1/2})}{d_n(x)}\biggr)^2 \cdot \frac{1}{d_n(x;K^{1/2})n}\sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f^{\ast}(X_i)\bigr)^2 K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr) \\
	& \leq \frac{(1 + o_P(1))\|\eta\|_{\infty}}{\alpha} \cdot \frac{1}{d_n(x;K^{1/2})n}\sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f^{\ast}(X_i)\bigr)^2 K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr),
	\end{align*}
	with the latter inequality following by Lemma~\ref{lem:measure_of_ball_estimate}. From here the proof proceeds along similar lines to the case where $\eta$ is non-negative.

	\textit{Proof of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0.5}.}  
	We will begin by introducing some notation. For a function $u \in L^2(\Omega)$, let
	\begin{equation*}
	L_{\varepsilon,n}u(x) := T_{\varepsilon,n}u(x) - u(x) = \frac{1}{d_n(x)n} \sum_{i = 1}^{n} \bigl(u(x) - u(x_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr)
	\end{equation*}
	and let
	\begin{equation*}
	\wt{L}_{\varepsilon,n}u(x) := \frac{d_n(x)}{d_P(x)}L_{\varepsilon,n}u(x) = \frac{1}{d_P(x)n} \sum_{i = 1}^{n} \bigl(u(x) - u(x_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
	\end{equation*}
	We have already shown that $d_n(x)/d_p(x) = (1 + o_P(1))$, and so we will focus our analysis on upper bounding $\Ebb[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2]$. We begin by providing expressions for the pointwise mean and variance of $\wt{L}_{\varepsilon,n}f^{\ast}(x)$:
	\begin{equation*}
	\Ebb\bigl[\wt{L}_{\varepsilon,n}f^{\ast}(x)\bigr] = T_{P,\varepsilon}f^{\ast}(x) - f^{\ast}(x)
	\end{equation*}
	and
	\begin{align*}
	\mathbb{V}\bigl[\wt{L}_{\varepsilon,n}f^{\ast}(x)\bigr] & = \frac{1}{(d_P(x))^2 n} \mathbb{V}\bigl[\bigl(f^{\ast}(X) - f^{\ast}(x)\bigr) \eta\bigl(\|X - x\|/\varepsilon\bigr)\bigr] \\
	& \leq \frac{1}{p_{\min}^2 \alpha^2 \varepsilon^{2d}} \mathbb{E}\bigl[\bigl(f^{\ast}(X) - f^{\ast}(x)\bigr)^2 K\bigl(\|X - x\|/\varepsilon\bigr)\bigr] \\
	& = \frac{1}{p_{\min}^2 \alpha^2 \varepsilon^{2d}} \int_{\Omega} \bigl(f^{\ast}(x') - f^{\ast}(x)\bigr)^2 K(\|x' - x\|/\varepsilon) p(x) \,dx.
	\end{align*}
	Then, integrating with respect to $P$ over $\Omega$, we obtain
	\begin{align*}
	\Ebb\bigl[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2\bigr]  & \leq \|T_{P,\varepsilon}f^{\star} - f^{\star}\|_{P}^2 + \frac{1}{p_{\min}^2 \alpha^2 \varepsilon^{2d}} E_{\varepsilon}(f^{\ast}) \\
	& \leq \|T_{P,\varepsilon}f^{\star} - f^{\star}\|_{P}^2 + \frac{C\varepsilon^2}{p_{\min}^2 \alpha^2 \varepsilon^{d}} |f|_{H^1(\Omega)}^2
	\end{align*}
	where $E_{\varepsilon}(f^{\ast}) := \int_{\Omega} \int_{\Omega} (f^{\ast}(x') - f^{\ast}(x))^2 K(\|x' - x\|/\varepsilon) p(x') p(x) ,\dx' \,dx$, and we have shown in our AISTATS paper that $E_{\varepsilon}(f^{\ast}) \leq C \varepsilon^{d + 2} \|f\|_{H^1(\Omega)}^2$, justifying the second inequality.
\end{proof}

>>>>

For a kernel function $\eta(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $\varepsilon > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{\varepsilon,Q}$ is given by
\begin{equation*}
\bigl(T_{\varepsilon,Q}f)(x) := \frac{1}{d_Q(x)} \int_{\Omega} f(z)\eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dQ(z)
\end{equation*}
where $d_Q(x) = \int_{\Omega} \eta\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$. We denote $T_{\varepsilon,P_n}$ by $T_{\varepsilon,n}$, and $d_{P_n}$ by $d_n$. For convenience, we will assume that $\eta$ is normalized, $\int_{0}^{\infty} \eta(s) \,ds = 1$, and we will denote $K(\cdot) = (\eta(\cdot))^2$.

The estimated function $\wh{f}$ we will consider is a kernel smoother passed over $\wh{\theta}$,
\begin{equation}
\label{eqn:kernel_smoother_laplacian_eigenmaps}
\wh{f}(x) := T_{\varepsilon,n}\wh{\theta}_{1:n}(x).
\end{equation}
In order to relate the out-of-sample error $\|\wh{f} - f^{\ast}\|_P$ to the in-sample error $\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n$, we will assume the following.
\begin{enumerate}[label=(A\arabic*)]
	\item
	\label{asmp:kernel}
	The kernel function $\eta$ is supported on a subset of $[0,1]$, with $\eta(1/2) > 0$. Additionally, $\eta$ is \textcolor{red}{Lipschitz} on its support.
	\item 
	\label{asmp:domain} There exists $\alpha > 0$ such that for every $x \in \mc{X}$, $d_{\nu}(x) \geq \alpha \cdot \varepsilon^d$.
	\item 
	\label{asmp:density}The distribution $P$ admits a density $p$ with respect to Lebesgue measure. The density $p \in C^1(\mc{X})$, and is bounded away from $0$ and $\infty$ for all $x \in \mc{X}$,
	\begin{equation*}
	0 < p_{\min} \leq p(x) < p_{\max} < \infty.
	\end{equation*}
\end{enumerate}
Under Assumption~\ref{asmp:kernel}, we automatically have that $d_\nu(x) = \varepsilon^d$ for all $x \in \mc{X}_{\varepsilon}$. Assumption~\ref{asmp:domain} is thus a condition on the boundary $\partial \Omega$, and is implied for all sufficiently small $\varepsilon$ if the boundary is Lipschitz. 
\begin{lemma}
	\label{lem:kernel_smoothing_laplacian_eigenmaps}
	Suppose $f^{\ast} \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$. Then with probability at least $1 - \delta - C\varepsilon^d\exp\{-Cn\varepsilon^d\}$, it holds that
	\begin{equation*}
	\|\wh{f} - f^{\ast}\|_P^2 \leq C\biggl(\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n^2 + \frac{1}{\delta} \cdot \frac{\varepsilon^2}{n\varepsilon^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2\biggr)
	\end{equation*}
\end{lemma}
Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data (on the order of $1/(n\varepsilon^d)$) by a factor of $\varepsilon^2$. The bias term, on the other hand, is standard, and in particular the following estimate of the bias can be obtained. 
\begin{lemma}
	Suppose $f^{\ast} \in H_0^{s}(\Omega)$, $p \in C^{r}(\Omega)$ for $r = s - 1$, and that $\eta$ is an order-$s$ kernel.
	\begin{equation*}
	\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2 \leq C \varepsilon^{2s}.
	\end{equation*}
\end{lemma}
\begin{proof}
	\textcolor{red}{(TODO)}
\end{proof}

@@

\begin{proof}[Proof of Lemma~\ref{lem:kernel_smoothing_laplacian_eigenmaps}]
	By the triangle inequality, 
	\begin{equation*}
	\|\wh{f} - f^{\ast}\|_P \leq \|\wh{f} - T_{\varepsilon,n}f^{\ast}_{1:n}\|_P + \|T_{\varepsilon,n}f^{\ast}_{1:n} - f^{\ast}\|_P.
	\end{equation*}
	We analyze each of the two terms in the above expression separately. We will show that
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_0}
	\|\wh{f} - T_{\varepsilon,n}f^{\ast}_{1:n}\|_P \leq \frac{1}{\alpha}\bigl(1 + o_P(1)\bigr) \cdot \|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n
	\end{equation}
	and 
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_0.5}
	\|T_{\varepsilon,n}f^{\ast}_{1:n} - f^{\ast}\|_P^2 \leq \frac{C}{\delta}\frac{\varepsilon^2}{n\varepsilon^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2,
	\end{equation}
	the latter inequality holding with probability $1 - \delta$.

@@

	The left hand side of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is equal to $\|T_{\varepsilon,n}(\wh{\theta}_{1:n} - f^{\ast}_{1:n})\|_P$, and~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is implied by showing that the operator $T_{\varepsilon,n}: L^2(X) \to L^2(\Omega)$ is (up to a constant) a contraction, meaning that for any $u \in L^2(X)$,
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_1}
	\|T_{\varepsilon,n}u\|_P^2 \leq \frac{1}{\alpha}\bigl(1 + o_P(1)\bigr) \|u\|_n^2.
	\end{equation}
	We begin with a pointwise estimate, which follows from Jensen's inequality:
	\begin{equation*}
	\Bigl(\bigl(T_{n,\varepsilon}u\bigr)(x)\Bigr)^2 \leq \frac{1}{d_n(x) n} \sum_{i = 1}^{n} \bigl(u(X_i)\bigr)^2 \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
	\end{equation*}
	(Note that it is here that we have used the positivity of $\eta(\cdot)$.) Integrating with respect to $P$ and exchanging sum with integral yields an upper bound on the $L^2(P)$ norm of $T_{n,\varepsilon}u$,
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_2}
	\|T_{n,\varepsilon}u\|_P^2 \leq \sum_{i = 1}^{n} u_i^2 \Bigl(\int_{\Omega} \frac{\eta(\|X_i - x\|/\varepsilon)}{n d_n(x)} p(x) \,dx \Bigr)
	\end{equation}
	The degree estimate $d_n(x)$ is quite close to $p(x)$. As we show in Lemma~\ref{lem:measure_of_ball_estimate}, if $\varepsilon = \omega((\log n/n)^{1/d})$ then the empirical measure $d_n(x)$ concentrates around $d_P(x)$ uniformly over $x \in \Omega$, 
	\begin{equation*}
	\sup_{x \in \Omega}|d_n(x) - d_P(x)| = o_P(\varepsilon^d);
	\end{equation*}
	moreover, it follows from Lipschitz continuity of $p$ that for any $x \in \Omega$,
	\begin{equation*}
	d_P(x) = \int_{\Omega} \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) p(x') \,dx' = p(x) d_{\nu}(x) + O(\varepsilon^{d + 1});
	\end{equation*}
	and thus by the triangle inequality $\sup_{x \in \Omega}|d_n(x) - d_\nu(x) p(x)| = o_P(\varepsilon^d)$.

@@

	Now, suppose $\eta(\cdot)$ can take on negative values. In this case, we have $\eta(\cdot) = \mathrm{sgn}(\eta(\cdot))K^{1/2}(\cdot)$. Let $d_n(x;K^{1/2})$ be the empirical degree with respect to kernel function $K^{1/2}$. Then
	\begin{align*}
	T_{n,\varepsilon}(u)(x) & = \frac{1}{d_n(x)n} \sum_{i = 1}^{n} \bigl(u(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr) \\
	& = \frac{d_n(x;K^{1/2})}{d_n(x)}\frac{1}{d_n(x;K^{1/2})n} \sum_{i = 1}^{n} \bigl(u(X_i)\bigr) \mathrm{sgn}\biggl(\eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr)\biggr) K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
	\end{align*}
	Taking the square and applying Jensen's inequality, we deduce that
	\begin{align*}
	\bigl(T_{n,\varepsilon}u(x)\bigr)^2 & \leq \biggl(\frac{d_n(x;K^{1/2})}{d_n(x)}\biggr)^2 \cdot \frac{1}{d_n(x;K^{1/2})n}\sum_{i = 1}^{n} \bigl(u(X_i)\bigr)^2 K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr) \\
	& \leq \frac{(1 + o_P(1))\|\eta\|_{\infty}^2}{\alpha^2} \cdot \frac{1}{d_n(x;K^{1/2})n}\sum_{i = 1}^{n} \bigl(u(X_i)\bigr)^2 K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr),
	\end{align*}
	with the latter inequality following by Lemma~\ref{lem:measure_of_ball_estimate} and assumption~\ref{asmp:kernel}. From here the proof proceeds along similar lines to the case where $\eta$ is non-negative.

	\textit{Proof of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0.5}.}  
	We will begin by introducing some notation. For a function $u \in L^2(\Omega)$, let
	\begin{equation*}
	L_{\varepsilon,n}u(x) := T_{\varepsilon,n}u(x) - u(x) = \frac{1}{d_n(x)n} \sum_{i = 1}^{n} \bigl(u(x) - u(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr)
	\end{equation*}
	and let
	\begin{equation*}
	\wt{L}_{\varepsilon,n}u(x) := \frac{d_n(x)}{d_P(x)}L_{\varepsilon,n}u(x) = \frac{1}{d_P(x)n} \sum_{i = 1}^{n} \bigl(u(x) - u(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
	\end{equation*}
	We have already shown that $d_n(x)/d_P(x) = (1 + o_P(1))$, and so we will focus our analysis on upper bounding $\Ebb[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2]$. We begin by providing expressions for the pointwise mean and variance of $\wt{L}_{\varepsilon,n}f^{\ast}(x)$:
	\begin{equation*}
	\Ebb\bigl[\wt{L}_{\varepsilon,n}f^{\ast}(x)\bigr] = T_{P,\varepsilon}f^{\ast}(x) - f^{\ast}(x)
	\end{equation*}
	and
	\begin{align*}
	\mathbb{V}\bigl[\wt{L}_{\varepsilon,n}f^{\ast}(x)\bigr] & = \frac{1}{(d_P(x))^2 n} \mathbb{V}\bigl[\bigl(f^{\ast}(X) - f^{\ast}(x)\bigr) \eta\bigl(\|X - x\|/\varepsilon\bigr)\bigr] \\
	& \leq \frac{1}{p_{\min}^2 \alpha^2 \varepsilon^{2d}} \mathbb{E}\bigl[\bigl(f^{\ast}(X) - f^{\ast}(x)\bigr)^2 K\bigl(\|X - x\|/\varepsilon\bigr)\bigr] \\
	& = \frac{1}{p_{\min}^2 \alpha^2 \varepsilon^{2d}} \int_{\Omega} \bigl(f^{\ast}(x') - f^{\ast}(x)\bigr)^2 K(\|x' - x\|/\varepsilon) p(x) \,dx.
	\end{align*}
	Then, integrating with respect to $P$ over $\Omega$, we obtain
	\begin{align*}
	\Ebb\bigl[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2\bigr]  & \leq \|T_{P,\varepsilon}f^{\star} - f^{\star}\|_{P}^2 + \frac{1}{p_{\min}^2 \alpha^2 \varepsilon^{2d}} E_{\varepsilon}(f^{\ast}) \\
	& \leq \|T_{P,\varepsilon}f^{\star} - f^{\star}\|_{P}^2 + \frac{C\varepsilon^2}{p_{\min}^2 \alpha^2 \varepsilon^{d}} |f|_{H^1(\Omega)}^2
	\end{align*}
	where $E_{\varepsilon}(f^{\ast}) := \int_{\Omega} \int_{\Omega} (f^{\ast}(x') - f^{\ast}(x))^2 K(\|x' - x\|/\varepsilon) p(x') p(x) ,\dx' \,dx$, and we have shown in our AISTATS paper that $E_{\varepsilon}(f^{\ast}) \leq C \varepsilon^{d + 2} |f|_{H^1(\Omega)}^2$, justifying the second inequality. Consequently, 
	\begin{align*}
	\|T_{n,\varepsilon}f^{\ast}_{1:n} - f^{\ast}\|_P^2 \leq (1 + o_P(1))^2 \|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2 & \leq \frac{\bigl(1 + o_P(1)\bigr)^2}{\delta} \mathbb{E}\bigl[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2\bigr] \\ 
	& \leq \frac{\bigl(1 + o_P(1)\bigr)^2}{\delta}\biggl(\|T_{P,\varepsilon}f^{\star} - f^{\star}\|_{P}^2 + \frac{C\varepsilon^2}{p_{\min}^2 \alpha^2 \varepsilon^{d}} |f|_{H^1(\Omega)}^2\biggr)
	\end{align*}
	with probability at least $1 - \delta$, establishing~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0.5}.
\end{proof}
